### 2024-12-03 paper/paper.qmd
Began to add in data section content, need to still make tables within this section

---
title: "..."
subtitle: "..."
author: 
  - Julia Lee
thanks: "Code and data are available at: https://github.com/jjlee-lee/Toronto_Neighbourhood_Income.git"
date: today
date-format: long
abstract: "..."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(tinytex)
library(knitr)
library(ggplot2)
library(dplyr)
library(arrow)
library(here)
library(modelsummary)
```

# Introduction

# Data {#sec-data}

To simulate, test, download, and clean the Neighbourhoods Profile data, the statistical programming language R was used [@citeR]. Specific libraries that assisted the analysis include `tidyverse` [@tidyverse], `dplyr` [@dplyr], `opendatatoronto` [@opendatatoronto], `tinytex` [@tinytex], `ggplot2` [@ggplot2], `knitr` [@knitr], `testthat` [@testthat], `here` [@here], `arrow` [@arrow], and `modelsummary` [@modelsummary].

## Neighbourhood Profiles Data

Neighbourhood Profiles data for the city of Toronto is provided by the Social Development, Finance & Administration from Toronto’s Open Data portal [@neighbourhoodprofile]. The dataset obtained from the city of Toronto’s Open Data portal [@opentorontodata], is a collection of data from the 2021 Canadian Census from Statistics Canada [@statisticscanada]. Records of socio-economic data for 158 geographic regions (i.e. social planning neighbourhoods) in Toronto can be found within this dataset. From age to ethnocultural diversity, the Census data highlights both the socio-economic and demographic characteristics of Toronto residents across its individual neighbourhoods.

The Neighbourhood Profiles data is an extensive dataset with each row representing a demographic or socio-economic characteristic and each column reflecting a Toronto neighbourhood. For each neighbourhood, there is information on its name, identification number, and status. Status refers to a neighbourhood’s Toronto Strong Neighbourhoods Strategy (TSNS) designation, which indicates whether a neighbourhood is an emerging area, an area that needs improvement, or neither.  Additionally, data for each neighbourhood’s age population, income, education, and more from the 2021 Census is provided. @tbl-rawdata below offers a preview of this dataset.

@tbl-rawdata displays neighbourhood names, numbers, and TSNS designations along with data for the number of individuals within different age groups in each neighbourhood. For example, Thistletown-Beaumond Heights is a neighbourhood in Toronto that needs improvement and has a fairly large number of children ages 0 to 14 years old. The rest of the dataset follows the same format, displaying different characteristics of all 158 neighbourhoods in the city of Toronto.

## Analysis Data

For the present analysis, the variables of interest are the different types of census family sizes along with the average after-tax income for each of Toronto’s neighbourhoods. The Neighbourhood Profiles data includes four different census family sizes: (1) 2-person families, (2) 3-person families, (3) 4-person families, and (4) five or more-person families. It also provides the average after-tax income ($) recorded in 2020 for each neighbourhood.

Since the objective is to find out what kinds of families are driving average neighbourhood income in Toronto, the data used throughout this analysis reflects the average income level for all 158 neighbourhoods alongside their counts of different census family sizes. With minimal data wrangling, the analysis data is simply an extraction of the larger dataset with a focus on the different family sizes and average after-tax income. @tbl-analysis below illustrates this analysis data for three neighbourhoods and summary statistics for this data can be found in the Appendix (@sec-appendix).

```{r}
#| echo: false
#| message: false
#| label: tbl-analysis
#| tbl-cap: "Toronto Neighbourhood Profiles Data"

analysis <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
#view(analysis)

for_table <- analysis[,c(1,2,4,5,6,7,10)]

for_table |>
  slice(1:3) |>
  kable(
    col.names = c("Name", "Number", "2 persons", "3 persons",
                  "4 persons", "5 or more persons", "Average Income"),
    digits = 1,
    linesep = "",
    align = c("l", "l", "c", "c", "c", "c", "c")
  )
```

In @tbl-analysis, the “Name” column represents each neighbourhood’s name, and the “Number” column reflects each neighbourhood’s identification number. The variables, “2 persons”, “3 persons”, “4 persons”, and “5 or more persons” represent the number of census families in each neighbourhood with those particular family sizes. Lastly, the “Average Income” variable reflects the average after-tax income for all neighbourhoods in 2020 (recorded in dollars). Within the analysis data file provided by this analysis, there are additional variables that represent the total number of census families by family size, the average census family size, and the average number of children in census families for each neighbourhood. These variables are considered to obtain additional context about how families are made up and distributed across Toronto’s neighbourhoods. It is important to note that while this additional information provided a better understanding of the variables of interest, they are not included in the model of the analysis. A detailed account of the model can be found in (@sec-model).

## Map Data

To further understand the type of families that drive average neighbourhood income, this analysis also uses Neighbourhoods data – a shapefile that contains geographic information of Toronto neighbourhood boundaries – to map the results of its model. The boundaries of each neighbourhood are defined using census tract information provided by Statistics Canada, and the shapefile itself is published by the Social Development, Finance & Administration from Toronto’s Open Data portal [@socialdevelopment]. The map within this analysis is created using ArcGIS Pro software [@arcgis]. By joining the analysis data shown in @tbl-analysis with the shapefile based on neighbourhood names, a map that highlights the distribution of family sizes that drive average neighbourhood income is built. The information contained in the shapefile is shown below in @tbl-mapdata.

The Neighbourhoods shapefile data contains neighbourhood names, numbers, and TSNS designations similar to the Neighbourhood Profiles data. It also includes geographic information (i.e. Shape_Length and Shape_Area) that defines the boundaries of each neighbourhood.

## A Note on Measurement

The Neighbourhood Profiles data is a reflection of the 2021 Census for neighbourhoods in Toronto, meaning that it measures and represents the demographic and socio-economic characteristics of families across Toronto at this particular time period. This information is collected through the use of questionnaires – a short and long-form census. The short-form census is sent out to all households, and it asks questions about simple characteristics like age, gender, and household size. With this information, the short-form census attempts to enumerate all individuals within a geographic region and may impute any missing data. The long-form census is sent out to only 25% of households and is a deeper source of data that outlines information such as housing, education, and ethnicity.

So, various aspects of families’ livelihoods are measured through the Census and aggregated to create an overall picture of particular geographic regions – in this case, Toronto neighbourhoods. It is important to note that through this measurement process, there is room for error as well as a need to be aware of this potential for error. Missing data values can be imputed and the aggregation of individual household characteristics or experiences can lead to incorrect assumptions about certain regions and their populations. Thus, this analysis presents its findings with this in mind.

# Model

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

## Model justification

## Model Validation

# Results

# Discussion

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Limiations & Future Directions

\newpage

\appendix

# Appendix {#sec-appendix}

## Analysis Data Summary Statistics

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-12-03 paper/paper.qmd
updated the README to reflect the addition of the shapefile, edited data section, added a reference, and rendered the pdf

@@
 To simulate, test, download, and clean the Neighbourhoods Profile data, the statistical programming language R was used [@citeR].
-Specific libraries that assisted the analysis include `tidyverse` [@tidyverse], `dplyr` [@dplyr], `opendatatoronto` [@opendatatoronto], `tinytex` [@tinytex], `ggplot2` [@ggplot2], `knitr` [@knitr], `testthat` [@testthat], `here` [@here], `arrow` [@arrow], and `modelsummary` [@modelsummary].
+Specific libraries that assisted the analysis include `tidyverse` [@tidyverse], `opendatatoronto` [@opendatatoronto], `tinytex` [@tinytex], `ggplot2` [@ggplot2], `knitr` [@knitr], `testthat` [@testthat], `here` [@here], `arrow` [@arrow], `modelsummary` [@modelsummary], and `sf` [@sf].
 
@@
 Additionally, data for each neighbourhood’s age population, income, education, and more from the 2021 Census is provided.
-@tbl-rawdata below offers a preview of this dataset.
+@tbl-rawdata below offers a small preview of this dataset.
 
-@tbl-rawdata displays neighbourhood names, numbers, and TSNS designations along with data for the number of individuals within different age groups in each neighbourhood.
-For example, Thistletown-Beaumond Heights is a neighbourhood in Toronto that needs improvement and has a fairly large number of children ages 0 to 14 years old.
-The rest of the dataset follows the same format, displaying different characteristics of all 158 neighbourhoods in the city of Toronto.
+```{r}
+#| echo: false
+#| message: false
+#| label: tbl-rawdata
+#| tbl-cap: "Toronto Neighbourhood Profiles Data"
+
+raw_data <- read_csv(here::here("data/01-raw_data/raw_data.csv"))
+#view(raw_data)
+
+for_show <- raw_data[,c(2,3)]
+#view(for_show)
+
+for_show |>
+  slice(1:6) |>
+  kable(
+    col.names = c("Neigbourhood Name", "West Humber-Clairville"),
+    digits = 1,
+    linesep = "",
+    align = c("l", "l")
+  )
+```
+
+@tbl-rawdata displays the name, number, and TSNS designation along with data for the number of individuals within different age groups for a neighbourhood (West Humber-Clairville) in Toronto.
+By looking at @tbl-rawdata, West Humber-Clairville is a neighbourhood that is not an emerging area and one that does not need improvement.
+It is also a region with a fairly large number of children ages 0 to 14 years old.
+The rest of the dataset follows the same format, displaying different characteristics (e.g.
+older age groups, education attainment, income levels, etc.) of all 158 neighbourhoods in the city of Toronto.
 
@@
 The Neighbourhood Profiles data includes four different census family sizes: (1) 2-person families, (2) 3-person families, (3) 4-person families, and (4) five or more-person families.
-It also provides the average after-tax income ($) recorded in 2020 for each neighbourhood.
+In simple terms, Statistics Canada defines a census family as one where all family members (related by blood marriage, common-law union, adoption, or a foster relationship) live together in the same dwelling [@censusdefinition].
+Census families can also be referred to as economic families [@economicdefinition].
+The data further provides the average after-tax income ($) recorded in 2020 for each neighbourhood.
 
@@
 With minimal data wrangling, the analysis data is simply an extraction of the larger dataset with a focus on the different family sizes and average after-tax income.
-@tbl-analysis below illustrates this analysis data for three neighbourhoods and summary statistics for this data can be found in the Appendix (@sec-appendix).
+@tbl-analysis below illustrates this analysis data for three neighbourhoods and summary statistics for this data can be found in the Appendix (@sec-appendixsummary).
 
@@
 #| label: tbl-analysis
-#| tbl-cap: "Toronto Neighbourhood Profiles Data"
+#| tbl-cap: "Toronto Neighbourhood Profiles Analysis Data"
 
@@
     linesep = "",
-    align = c("l", "l", "c", "c", "c", "c", "c")
+    align = c("l", "c", "c", "c", "c", "c", "c")
   )
@@
 The map within this analysis is created using ArcGIS Pro software [@arcgis].
-By joining the analysis data shown in @tbl-analysis with the shapefile based on neighbourhood names, a map that highlights the distribution of family sizes that drive average neighbourhood income is built.
+By joining the analysis data shown in @tbl-analysis with the shapefile based on neighbourhood names, a map that highlights the distribution of family sizes that drive average neighbourhood income can be created.
 The information contained in the shapefile is shown below in @tbl-mapdata.
 
-The Neighbourhoods shapefile data contains neighbourhood names, numbers, and TSNS designations similar to the Neighbourhood Profiles data.
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
+#| label: tbl-mapdata
+#| tbl-cap: "Toronto Neighbourhood Location & Boundaries Data"
+
+library(sf)
+
+map_data <- st_read("~/Toronto_Neighbourhood_Income/data/03-map_data", quiet=TRUE)
+#view(map_data)
+
+for_illustration <- map_data[,c(1,6,7,9)]
+#view(for_illustration)
+
+for_illustration |>
+  slice(1:3) |>
+  kable(
+    col.names = c("Object ID", "Neighbourhood Number", "Neighbourhood Name", 
+                  "TSNS Designation", "Geometry"),
+    digits = 1,
+    booktabs = TRUE,
+    linesep = "",
+    align = c("l", "l", "l", "l", "l"),
+    format.args = list(big.mark = ",")
+  )
+```
+
+The Neighbourhoods shapefile data contains neighbourhood names, numbers, and TSNS designations similar to the Neighbourhood Profiles data (@tbl-rawdata).
 It also includes geographic information (i.e.
-Shape_Length and Shape_Area) that defines the boundaries of each neighbourhood.
+the geometry/coordinates of a neighbourhood) that defines the boundaries of each neighbourhood as shown in @tbl-mapdata.
 
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-## Analysis Data Summary Statistics
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+## Analysis Data Summary Statistics {#sec-appendixsummary}
 
@@
 #| echo: false
-#| eval: true
 #| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+#| label: tbl-summarystatistics
+#| tbl-cap: "Toronto Neighbourhood Profiles Data Summary Statistics"
 
-plot(first_model, "trace")
+summary_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
+#view(summary_data)
 
-plot(first_model, "rhat")
+for_summary <- summary_data[,c(4,5,6,7,10)]
+#view(for_summary)
+
+# Turn values into numeric data type
+
+for_summary$`2 persons` = as.numeric(as.character(for_summary$`2 persons`))
+for_summary$`3 persons` = as.numeric(as.character(for_summary$`3 persons`))
+for_summary$`4 persons` = as.numeric(as.character(for_summary$`4 persons`))
+for_summary$`5 or more persons` = as.numeric(as.character(for_summary$`5 or more persons`))
+for_summary$`Average after-tax income of economic family in 2020 ($)` = as.numeric(as.character(for_summary$`Average after-tax income of economic family in 2020 ($)`))
+
+final_summary <- summary(for_summary)
+#view(final_summary)
+
+final_summary |>
+    kable(
+    col.names = c("2-person Families", "3-person Families", "4-person Families",
+                  "5 or more-person Families", "Average Income"),
+    digits = 1,
+    booktabs = TRUE,
+    linesep = "",
+    align = c("l", "l", "l", "l", "l"),
+  )
 ```
+
+@tbl-summarystatistics presents the summary statistics for each census family size and the average after-tax income across all 158 neighbourhoods in Toronto.
 

### 2024-12-03 scripts/05-model_data.R
removed unnecessary library from model script

@@ -12,7 +12,6 @@
 #### Setting Up the Workspace ####
 
 library(tidyverse)
-library(testthat)
 library(arrow)
 library(here)
 library(dplyr)

### 2024-12-03 paper/paper.qmd
Completed model section along with its corresponding appendix and rendered the pdf

@@
 
-## Model set-up
+To understand the types of families that drive average neighbourhood income, this analysis constructs a multiple linear regression model using R [@citeR] that considers average neighbourhood income as the response variable and the different census family sizes as predictors.
+As mentioned in @sec-data, the average neighbourhood income variable reflects the average after-tax income of all census families for each neighbourhood in 2020 (recorded in dollars).
+Different family sizes (i.e.
+2 to 5 or more-person families) refer the number of family members that live in the same dwelling within each individual neighbourhood.
+With this, the final model within this analysis is structured as follows:
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+$$
+\hat{y} = b_0 + b_1 \cdot x_1 + b_2 \cdot x_2 + b_3 \cdot x_3 + \epsilon
+$$
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+where
+
+-   $\hat{y}$ represents the average neighbourhood after-tax income,
+
+-   $b_0$ represents the intercept of the regression model,
+
+-   $b_1$ represents the effect of two-person census families,
+
+-   $x_1$ represents the number of two-person census families,
+
+-   $b_2$ represents the effect of three-person census families,
+
+-   $x_2$ represents the number of three-person census families,
+
+-   $b_3$ represents the effect of four-person census families,
+
+-   $x_3$ represents the number of four-person census families,
+
+-   $\epsilon$ captures the error within the regression model
+
+It is important to highlight that as a result of model selection and individual predictor t-tests, families with 5 or more members are not considered in the final model.
+The impact of 2, 3, and 4-person families on average neighbourhood income across Toronto neighbourhoods are examined through this model.
 
@@
 
-## Model Validation
+As the objective of the analysis is to understand what might be driving the average neighbourhood income in terms of family size, a multiple linear regression will allow for an examination of the relationship between different census family sizes and average income.
+By striving to preserve the interpretability of the model, this analysis can further determine which family sizes are meaningful influencers of a neighbourhood’s average after-tax income.
 
-# Results
+The underlying assumption of linearity that multiple linear regression models hold can be a potential limitation as these models will not be able accurately explain the influence of predictors on a response variable if these variables have a non-linear relationship with each other.
+The use of a multiple linear regression may not be the best choice in these cases.
+However, with average neighbourhood income and increasing census family sizes, the model’s diagnostic plots provide evidence for linearity, meaning that the use of a multiple linear regression model can be suitable for this analysis.
+Further justification and validation for the model can be found in the Appendix (@sec-appendixmodel).
 
-# Discussion
+# Results {#sec-results}
 
-## First discussion point {#sec-first-point}
-
-## Second discussion point
-
-## Third discussion point
+# Discussion {#sec-discussion}
 
@@
 
+## Model Validation {#sec-appendixmodel}
+
+The construction of a model for this analysis began by fitting an initial model with average neighbourhood income as the response variable and 2, 3, 4, and 5 or more-person families as predictors.
+As seen by @tbl-initialmodel and the individual t-tests, 5 or more-person families do not appear to significantly impact a neighbourhood’s average income with the largest p-value among all predictors (p > 0.001).
+
+```{r}
+#| warning: false
+#| echo: false
+#| label: tbl-initialmodel
+#| tbl-cap: "Summary Table of the Initial Model"
+
+model_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
+#view(model_data)
+
+# Fitting the initial regression model
+
+## Changing column names
+
+regression_data <-
+  model_data |>
+  rename(
+    neighbourhood_name = "Neighbourhood Name",
+    neighbourhood_num = "Neighbourhood Number",
+    num_census_fam = "Total - Census families in private households by family size - 25% sample data",
+    size_two = "2 persons",
+    size_three = "3 persons",
+    size_four = "4 persons",
+    size_five_plus = "5 or more persons",
+    avg_fam_size = "Average size of census families",
+    avg_num_child = "Average number of children in census families with children",
+    avg_income = "Average after-tax income of economic family in 2020 ($)"
+  )
+
+#view(regression_data)
+
+## Turn columns into numerical vectors
+## Hints for the code of this step were provided by:
+## https://stackoverflow.com/questions/61260686/too-many-coefficients-with-lm
+
+regression_data$size_two = as.numeric(as.character(regression_data$size_two))
+regression_data$size_three = as.numeric(as.character(regression_data$size_three))
+regression_data$size_four = as.numeric(as.character(regression_data$size_four))
+regression_data$size_five_plus = as.numeric(as.character(regression_data$size_five_plus))
+regression_data$avg_income = as.numeric(as.character(regression_data$avg_income))
+
+initial_model <- lm(avg_income ~ size_two + size_three + size_four + 
+                      size_five_plus, data = regression_data)
+
+modelsummary(initial_model, statistic = "{std.error} ({p.value})")
+```
+
+So, another model was created without this variable (i.e. a reduced model), and all predictors seem to have significant impacts (p < 0.001) on neighbourhood income as shown in @tbl-newmodel.
+
+```{r}
+#| warning: false
+#| echo: false
+#| label: tbl-newmodel
+#| tbl-cap: "Summary Table of the Reduced Model"
+
+model <- readRDS(here("model/model_1.rds"))
+
+modelsummary(model, statistic = "{std.error} ({p.value})")
+```
+
+Model selection was further conducted by comparing the BIC scores of both models (shown in @tbl-bic).
+As the new model has a lower BIC score, this is the one that is used within the analysis and outlined in @sec-model.
+
+```{r}
+#| echo: false
+#| message: false
+#| label: tbl-bic
+#| tbl-cap: "BIC Scores for Initial and Reduced Models"
+
+bic_scores <-
+  tibble(
+    Model = c("Initial", "Reduced"),
+    BIC = c(BIC(initial_model), BIC(model))
+  )
+
+#view(bic_scores)
+
+bic_scores |>
+    kable(
+    col.names = c("Model", "BIC Score"),
+    digits = 1,
+    booktabs = TRUE,
+    linesep = "",
+    align = c("c", "c"),
+)
+```
+
+The purpose of the model is to be able to interpret the coefficient estimates and what they mean in terms of influencing average neighbourhood income.
+For this reason, transformations (e.g.
+log transformations on predictor variables) were not applied to the data even with the presence of some evidence against the linear regression model assumptions.
+
+Looking at the diagnostic plots (@fig-residualsandpredictors), the lack of observational patterns in the Residuals vs Predictor plots for 2, 3, and 4-person families points to evidence for the linearity assumption of linear regression models.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-residualsandpredictors
+#| fig-cap: Showing the Residual vs. Predictors
+
+residuals_1 <- model$residuals
+#summary(residuals_1)
+
+par(mfrow=c(1,3))
+plot(residuals_1 ~ regression_data$size_two, main = "Residuals vs 2-person Families", 
+     xlab = "Number of 2-person Families", ylab = "Residuals")
+plot(residuals_1 ~ regression_data$size_three, main = "Residuals vs 3-person Families", 
+     xlab = "Number of 3-person Families", ylab = "Residuals")
+plot(residuals_1 ~ regression_data$size_four, main = "Residuals vs 4-person Families", 
+     xlab = "Number of 4-person Families", ylab = "Residuals")
+```
+The Residual vs Fitted Values plot (@fig-residualsandfitted) seems to have some clustering of points, indicating that the constant variance and independence assumptions are somewhat satisfied.
+However, because the overall pattern of the plot is still relatively spread out with no observable patterns, and as the objective of the model is interpretability, transformations were not applied to the variables in the model to better these assumptions.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-residualsandfitted
+#| fig-cap: Showing the Residual vs. Fitted Values
+
+fitted_1 <- model$fitted.values
+#summary(fitted_1)
+
+plot(residuals_1 ~ fitted_1, main = "Residuals vs Fitted Values", 
+     xlab = "Fitted Values", ylab = "Residuals")
+abline(h = 0, col ="red")
+```
+Lastly, though the Normal Q-Q plot (@fig-normal) shows some deviation from the Q-Q line, the normality assumption is somewhat satisfied with little evidence to suggest an alarming violation of this assumption.
+Putting all this together, while improvements can be made to the model for better, more accurate predictions, it is sufficient for the purpose of interpretability.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-normal
+#| fig-cap: Showing the Normal Q-Q Plot
+
+qqnorm(residuals_1)
+qqline(residuals_1)
+```
+
 \newpage

### 2024-12-03 paper/paper.qmd
working on creating a figure for results section

@@
 
+```{r}
+#| echo: false 
+#| message: false
+#| label: fig-familysizedistribution
+#| fig-cap: Number of 2, 3, and 4-person Census Families Across Toronto Neighbourhoods
+
+family_size <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
+#view(family_size)
+
+# Turn values into numeric data type
+
+family_size$`Neighbourhood Number` = as.numeric(as.character(family_size$`Neighbourhood Number`))
+family_size$`2 persons` = as.numeric(as.character(family_size$`2 persons`))
+family_size$`3 persons` = as.numeric(as.character(family_size$`3 persons`))
+family_size$`4 persons` = as.numeric(as.character(family_size$`4 persons`))
+
+ggplot(family_size, aes(x = "Neighbourhood Name", y = "2 persons")) + 
+  geom_line(linewidth = 0.5, colour="black") + geom_point(colour = "black") +
+ theme_light()  
+  ggtitle("EMS Demand From 2017 to 2022") +
+  labs(x = "Year", y = "Demand for EMS", caption = "Data Source: Toronto Paramedic Services") + 
+  theme(plot.title = element_text(hjust = 0.5)) 
+```
+
 # Discussion {#sec-discussion}

### 2024-12-04 paper/paper.qmd
finished results section, added a new folder in "plan" with maps, and rendered pdf

--- 2024-12-03 paper/paper.qmd
@@
 Census families can also be referred to as economic families [@economicdefinition].
-The data further provides the average after-tax income ($) recorded in 2020 for each neighbourhood.
+The data further provides the average after-tax income (\$) recorded in 2020 for each neighbourhood.
 
@@
 
+## Distribution of Different Census Family Sizes Across Toronto Neighbourhoods
+
+```{r}
+#| echo: false 
+#| message: false
+#| label: fig-familysizetwo
+#| fig-cap: Number of 2-person Census Families Across Toronto Neighbourhoods
+
+family_size <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
+#view(family_size)
+
+# Rename columns
+
+fam_size_data <-
+  family_size |>
+  rename(
+    neighbourhood_name = "Neighbourhood Name",
+    neighbourhood_num = "Neighbourhood Number",
+    num_census_fam = "Total - Census families in private households by family size - 25% sample data",
+    size_two = "2 persons",
+    size_three = "3 persons",
+    size_four = "4 persons",
+    size_five_plus = "5 or more persons",
+    avg_fam_size = "Average size of census families",
+    avg_num_child = "Average number of children in census families with children",
+    avg_income = "Average after-tax income of economic family in 2020 ($)"
+  )
+
+# Turn values into numeric data type
+
+fam_size_data$size_two = as.numeric(as.character(fam_size_data$size_two))
+fam_size_data$size_three = as.numeric(as.character(fam_size_data$size_three))
+fam_size_data$size_four = as.numeric(as.character(fam_size_data$size_four))
+
+ggplot(fam_size_data, aes(x = neighbourhood_num, y = size_two)) + 
+  geom_bar(stat = "identity", colour = "blue", fill = "lightblue") + theme_light() +  
+  ggtitle("2-person Families in Toronto Neighbourhoods") +
+  labs(x = "158 Toronto Neighbourhoods", y = "Number of 2-person Families", 
+       caption = "Data Source: Statistics Canada, 2021 Census") + 
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(axis.text.x = element_blank()
+  )
+```
+
+```{r}
+#| echo: false 
+#| message: false
+#| label: fig-familysizethree
+#| fig-cap: Number of 3-person Census Families Across Toronto Neighbourhoods
+
+ggplot(fam_size_data, aes(x = neighbourhood_num, y = size_three)) + 
+  geom_bar(stat = "identity", colour = "purple", fill = "pink") + theme_light() +  
+  ggtitle("3-person Families in Toronto Neighbourhoods") +
+  labs(x = "158 Toronto Neighbourhoods", y = "Number of 3-person Families", 
+       caption = "Data Source: Statistics Canada, 2021 Census") + 
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(axis.text.x = element_blank()
+  )
+```
+
+```{r}
+#| echo: false 
+#| message: false
+#| label: fig-familysizefour
+#| fig-cap: Number of 4-person Census Families Across Toronto Neighbourhoods
+
+ggplot(fam_size_data, aes(x = neighbourhood_num, y = size_four)) + 
+  geom_bar(stat = "identity", colour = "darkgreen", fill = "lightgreen") + theme_light() +  
+  ggtitle("4-person Families in Toronto Neigbourhoods") +
+  labs(x = "158 Toronto Neighbourhoods", y = "Number of 4-person Families", 
+       caption = "Data Source: Statistics Canada, 2021 Census") + 
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme(axis.text.x = element_blank()
+  )
+```
+
+@fig-familysizetwo, @fig-familysizethree, and @fig-familysizefour illustrate the number of 2, 3, and 4-person families, respectively, across Toronto’s 158 neighbourhoods.
+From this, 2-person census families appear to be the most common as almost all neighbourhoods have close to 1000 2-person families with several of them having over 3000, near 4000 2-person families.
+On the other hand, 3 and 4-person families seem to less common with more neighbourhoods having less than 1000 families of these sizes (i.e.
+under or just over 500 families) and with the maximum number of 3 or 4-person families being just under 3000.
+These observations seem to suggest that given the higher numbers of 2-person families across Toronto neighbourhoods, this family size may be a key driver of average neighbourhood income within the city of Toronto.
+
+## Model Results
+
+```{r}
+#| warning: false
+#| echo: false
+#| label: tbl-modelresults
+#| tbl-cap: "Summary Results of Analysis Model"
+
+model_results <- readRDS(here("model/model_1.rds"))
+
+modelsummary(model_results, statistic = "{std.error} ({p.value})")
+```
+
+The results from the model shown in @tbl-modelresults indicate significant relationships between average neighbourhood income and different census family sizes (p \< 0.001 for each family size).
+This means that the addition of a 2-person census family to a neighbourhood is expected to increase the neighbourhood’s average after-tax income by around \$28.59 (± 4.76).
+Likewise, the addition of a 4-person family to a neighbourhood is expected to increase the neighbourhood’s average after-tax income by approximately \$148.40 (± 16.32).
+Interestingly, the addition of a 3-person census family to a neighbourhood is expected to decrease the neighbourhood’s average after-tax income by around \$184.42 (± 17.66).
+These findings seem to suggest that 4-person families are also be key drivers of average neighbourhood income.
+Additionally, this may imply that on average, while 2 and 4-person families can commonly have above-average incomes, more 3-person families may have below-average incomes across Toronto’s neighbourhoods.
+
+## Map Visualization
+
+![Distribution of 4-person Census Families Across Toronto](images/map_visualization_four-02.png){#fig-map}
+
+![Distribution of 2-person Census Families Across Toronto](images/map_visualization_two-02.png){#fig-map2}
+
+The initial analysis of the distribution of different census family sizes across Toronto neighbourhoods highlighted that 2-person families can be key drivers of average neighbourhood income.
+The results of the model further suggest that 2 and 4-person families positively contribute to average neighbourhood income.
+With this in mind, @fig-map maps the distribution of 4-person census families throughout the city of Toronto.
+This shows how the majority of 4-person census families are found in the neighbourhoods towards the east end of the city with a few neighbourhoods near the city’s centre and the Harbourfront area.
+@fig-map2 demonstrates the distribution of 2-person families.
+Here, there appears to be a shift in that more 2-person census families seem to be in neighbourhoods closer to the city’s centre and the west side of the Harbourfront area.
+These illustrations provide a better understanding of where the key drivers of average neighbourhood income are located.
+
 # Discussion {#sec-discussion}
@@
 The construction of a model for this analysis began by fitting an initial model with average neighbourhood income as the response variable and 2, 3, 4, and 5 or more-person families as predictors.
-As seen by @tbl-initialmodel and the individual t-tests, 5 or more-person families do not appear to significantly impact a neighbourhood’s average income with the largest p-value among all predictors (p > 0.001).
+As seen by @tbl-initialmodel and the individual t-tests, 5 or more-person families do not appear to significantly impact a neighbourhood’s average income with the largest p-value among all predictors (p \> 0.001).
 
@@
 
-So, another model was created without this variable (i.e. a reduced model), and all predictors seem to have significant impacts (p < 0.001) on neighbourhood income as shown in @tbl-newmodel.
+So, another model was created without this variable (i.e. a reduced model), and all predictors seem to have significant impacts (p \< 0.001) on neighbourhood income as shown in @tbl-newmodel.
 
@@
 ```
+
 The Residual vs Fitted Values plot (@fig-residualsandfitted) seems to have some clustering of points, indicating that the constant variance and independence assumptions are somewhat satisfied.
@@
 ```
+
 Lastly, though the Normal Q-Q plot (@fig-normal) shows some deviation from the Q-Q line, the normality assumption is somewhat satisfied with little evidence to suggest an alarming violation of this assumption.

### 2024-12-04 paper/paper.qmd
updated paper quarto document with discussion section and rendered the pdf

@@
 
-## A Note on Measurement
+## A Note on Measurement {#sec-datameasure}
 
@@
 
+## Summary
+
+The present analysis aims to understand the kinds of families that drive average neighbourhood income in Toronto.
+To learn more about the families that reside in Toronto neighbourhoods, the analysis observes the distribution of different census family sizes with @fig-familysizetwo, @fig-familysizethree, and @fig-familysizefour.
+It also illustrates the geographic locations of Toronto families with @fig-map and @fig-map2.
+By constructing a multiple linear regression model that examines the relationship between average neighbourhood income and different census family sizes, the analysis provides further support for its observations and findings.
+
+## Analysis Findings
+
+A study like the present analysis can help support planning decisions for individual neighbourhoods along with local and general businesses.
+The analysis finds that 2 and 4-person census families are key drivers of average neighbourhood income.
+
+Considering the results of the model, the output shows that the baseline average neighbourhood income is approximately $125, 318 – meaning that Toronto neighbourhoods may have a healthy level of disposable income.
+This opens up opportunities for neighbourhoods as well as businesses.
+The question is, however, which neighbourhoods are more affluent and how can neighbourhoods raise the income level to become economically stronger?
+The answer to this can be found by looking at the impact of different census family sizes.
+To start, 2-person census families drive the average neighbourhood income by around $28 per family whereas 4-person homes drive income by approximately $148.
+While the amount of income that 2-person families contribute is smaller than that of 4-person families, the impact that 2-person families have on neighbourhood income can be just as large because this family size is most commonly found across Toronto’s neighbourhoods.
+Further, 3-person families are found to decrease the income by around $184, suggesting that these families may have lower incomes compared to other family sizes.
+
+With this, organizations could use this information to make planning decisions.
+For instance, neighbourhoods could target 2 or 4-person census families when planning for housing in the interest of attracting families that will stimulate the economy.
+Additionally, local and general businesses could focus on these even-numbered families as a means of effectively promoting their products or shaping their brand.
+This is not to say that 3-person census families should be ignored.
+However, it should be noted that for every 3-person family, it takes approximately one 2-person family plus a 4-person family to make up for the decrease in average neighbourhood income.
+The impact of 3-person families on neighbourhood could also be mitigated by reason of the large number of 2-person families that live throughout Toronto.
+In short, for planning decisions in the interest of increasing a neighbourhood or business’ income position, there may be a benefit in considering the interests or need of 2 and 4-person families.
+
+This analysis further visualizes the neighbourhoods where the most 2 and 4-person families live.
+Knowing the geographic regions in which the key drivers of average neighbourhood income are located can further help work towards a more equitable distribution of opportunities and resources.
+For example, in the neighbourhoods that have many 2 or 4-person census families, more opportunities such as higher quality schools or higher employment availability may be present and more resources could be allocated to these neighbourhoods to further support or maintain those opportunities.
+On the other hand, understanding where the drivers of neighbourhood income are also helps to identify the areas where there are more families such as 3-person families who are not able to contribute as much to the average neighbourhood income.
+By using this information, planning decisions could be made to provide these neighbourhoods with resources for better opportunities with education and employment, for example, to ensure equitable resource distributions across all Toronto neighbourhoods.
+
 ## Limiations & Future Directions
+
+Even with the deeper understanding that this analysis brings forth, there are some limitations to consider.
+The model within this analysis centers around interpretability, meaning that despite seeing some evidence against its underlying assumptions, transformations were not applied to the data to resolve these diagnostic issues.
+As the application of transformations make it challenging to interpret coefficient estimates and what they mean in the context of the analysis, the present analysis did not engage in this step of the model building process.
+However, this limits the model’s ability to make more accurate predictions that capture what the average neighbourhood income would be with the addition of 2, 3, or 4-person census families to a given neighbourhood.
+So, findings that point to exact numbers of average neighbourhood income for different numbers of varying family sizes could not be included in the analysis.
+
+In addition to this, the potential for error that exists within the Neighbourhood Profiles data from the 2021 Canadian Census (as briefly mentioned in @sec-datameasure) can be another limitation that hints at a compelling future direction for this analysis.
+As the Neighbourhood Profiles data is the aggregation of individual family or household characteristics and experiences, it can provide a better understanding of the overall picture – that is, the general patterns and trends within the city of Toronto.
+However, it is limited in capturing unique or individual characteristics which can be lost during its process of aggregating data.
+Thus, with this overall understanding about the key drivers of average neighbourhood income, future research could be focused on individual neighbourhoods (the neighbourhoods with many 4-person families, for instance).
+These future studies could look at how key families interact with the local resources (e.g.
+grocery stores, health services, etc.) available to them to understand how they contribute to their neighbourhood’s average income.
+It may also be interesting to examine other characteristics (e.g.
+education attainment, ethnicity, etc.) that might explain why 2 and 4-person families are drivers of the average income for their neighbourhoods.
 
@@
 
+\newpage
+
 # References

### 2024-12-04 paper/paper.qmd
updated README to reflect the new "images" folder in the "paper" folder, added abstract, and edited figures

@@
 date-format: long
-abstract: "..."
+abstract: "By examining Toronto Neighbourhood Profiles data, the following analysis aims to understand what kinds of families drive average neighbourhood income to help support planning decisions for individual neighbourhoods along with local and general businesses.
+With both data visualization and the application of a linear regression model, this analysis finds that 2-person and 4-person families are key drivers of average neighbourhood income throughout Toronto.
+This implies that focusing neighbourhood or business planning decisions on these types of families may help to increase a neighbourhood or business’ income position.
+Further, this analysis highlights that additional considerations for the unique characteristics of these key driver families are important to understand the interplay between families and neighbourhood income in more depth."
 format: pdf
@@
 # Introduction
+
+The interactions families have with their surrounding environments (i.e.
+their neighbourhoods) can be influenced by both the characteristics of the environment and the kinds of resources that families have.
+Income and financial resources can be strong determinants of the types of opportunities (e.g.
+housing, education, employment, etc.) families are able to access within their neighbourhoods [@resource].
+With this, to better understand the concept of income across the city of Toronto’s neighbourhoods, this analysis aims to identify and determine the kinds of families that are driving average neighbourhood income levels.
+
+This analysis can help support and inform planning decisions for individual neighbourhoods along with local and general businesses.
+With a better understanding of the livelihoods of families in Toronto, it may be possible to better allocate resources to the neighbourhoods that have few families that drive the average neighbourhood income.
+Further, planning decisions that center around effectively targeting or considering the families that drive neighbourhood income may help to strengthen the income position of various neighbourhoods or local businesses.
+Thus, this analysis serves as a starting point to further understand the interaction between families and neighbourhood income.
+
+By considering data from the 2021 Canadian Census, this analysis begins by investigating the distribution of different census family sizes (i.e.
+2, 3, or 4-person families) across all 158 of Toronto’s neighbourhoods.
+Next, it creates a multiple linear regression model that examines the relationship between average neighbourhood income and different family sizes.
+The results of this model demonstrate the family sizes that significantly impact neighbourhood income.
+Lastly, after identifying key drivers of average neighbourhood income through the model, the analysis considers the geographic locations of those drivers.
+As a result, the present analysis finds that 2 and 4-person families are key drivers of average neighbourhood income.
+Additionally, as there are varying numbers of 2 and 4-person families across Toronto neighbourhoods, it is important to also consider the characteristics of individual families to ensure effective neighbourhood or local business planning decisions.
+
+The following provides an in-depth description of the data used within this analysis [@sec-data], an explanation of the analysis model [@sec-model], a detailed account of the findings [@sec-results], and a thorough discussion of the overall analysis [@sec-discussion].
 
@@
   geom_bar(stat = "identity", colour = "blue", fill = "lightblue") + theme_light() +
-  ggtitle("2-person Families in Toronto Neighbourhoods") +
-  labs(x = "158 Toronto Neighbourhoods", y = "Number of 2-person Families", 
+  labs(x = "Toronto Neighbourhoods (158)", y = "Number of 2-person Families", 
        caption = "Data Source: Statistics Canada, 2021 Census") + 
@@
   geom_bar(stat = "identity", colour = "purple", fill = "pink") + theme_light() +
-  ggtitle("3-person Families in Toronto Neighbourhoods") +
-  labs(x = "158 Toronto Neighbourhoods", y = "Number of 3-person Families", 
+  labs(x = "Toronto Neighbourhoods (158)", y = "Number of 3-person Families", 
        caption = "Data Source: Statistics Canada, 2021 Census") + 
@@
   geom_bar(stat = "identity", colour = "darkgreen", fill = "lightgreen") + theme_light() +
-  ggtitle("4-person Families in Toronto Neigbourhoods") +
-  labs(x = "158 Toronto Neighbourhoods", y = "Number of 4-person Families", 
+  labs(x = "Toronto Neighbourhoods (158)", y = "Number of 4-person Families", 
        caption = "Data Source: Statistics Canada, 2021 Census") + 
@@
 par(mfrow=c(1,3))
-plot(residuals_1 ~ regression_data$size_two, main = "Residuals vs 2-person Families", 
+plot(residuals_1 ~ regression_data$size_two, 
      xlab = "Number of 2-person Families", ylab = "Residuals")
-plot(residuals_1 ~ regression_data$size_three, main = "Residuals vs 3-person Families", 
+plot(residuals_1 ~ regression_data$size_three, 
      xlab = "Number of 3-person Families", ylab = "Residuals")
-plot(residuals_1 ~ regression_data$size_four, main = "Residuals vs 4-person Families", 
+plot(residuals_1 ~ regression_data$size_four, 
      xlab = "Number of 4-person Families", ylab = "Residuals")
@@
 
-plot(residuals_1 ~ fitted_1, main = "Residuals vs Fitted Values", 
+plot(residuals_1 ~ fitted_1, 
      xlab = "Fitted Values", ylab = "Residuals")

### 2024-12-04 paper/paper.qmd
fixed format of overall pdf
updated paper with title and subtitle

@@
 ---
-title: "..."
-subtitle: "..."
+title: "How do families shape average neighbourhood incomes in Toronto?"
+subtitle: "Different family sizes appear to impact income in different ways with 2 and 4-person families being key drivers of average neighbourhood incomes in Toronto"
 author: 
@@
 With this, to better understand the concept of income across the city of Toronto’s neighbourhoods, this analysis aims to identify and determine the kinds of families that are driving average neighbourhood income levels.
+
+Following this, the variable of interest or the estimand within this analysis is the average neighbourhood income.
+As income can influence the kinds of opportunities and resources families in various neighbourhoods across Toronto have access to, this analysis seeks to better understand how neighbourhood average income levels are shaped by considering the impacts of different family sizes.
 
@@
 
+\newpage
+
 ## Model Results
@@
 Additionally, this may imply that on average, while 2 and 4-person families can commonly have above-average incomes, more 3-person families may have below-average incomes across Toronto’s neighbourhoods.
+
+\newpage
 

### 2024-12-04 paper/paper.qmd
Added another section to the appendix to complete the paper and rendered the pdf

@@
 education attainment, ethnicity, etc.) that might explain why 2 and 4-person families are drivers of the average income for their neighbourhoods.
+A discussion of how this research might be conducted using observational data and surveys can be found in @sec-appendixobs.
 
@@
 
+## Using observational data and surveys to better understand family characteristics at the neighbourhood level {#sec-appendixobs}
+
+As discussed in @sec-discussion, a compelling future direction is to look at how key families interact with the local resources (e.g.
+grocery stores, health services, etc.) available to them to understand how they contribute to their neighbourhood’s average income.
+The following outlines a research plan for this direction.
+
+To explore this new research direction, a participant observation and survey method can be used to recognize how families of different sizes interact with local resources and contribute to their neighbourhood’s average income.
+The participant observation will take place at a specific neighbourhood in Toronto and provide a clearer idea of how 2 or 4-person families spend their income (i.e.
+what neighbourhood resources they spend their income on).
+To compliment this, a survey will be distributed to all families in a particular neighbourhood to gain a better understanding of other characteristics such as education attainment or ethnicity families within that neighbourhood have.
+The survey can also highlight more about the reason behind family’s decisions and the behaviours observed.
+
+Field notes that contain descriptions of family’s behaviour such as how often 4-person families visit a grocery store or a library will be taken and collected throughout this observation process.
+In terms of the survey, its structure and the kinds of questions it contains will depend on the neighbourhood.
+If the analysis is taking place at a neighbourhood with a higher number of 4-person families, then it is likely that there more families with children.
+So, questions could be tailored to their experiences such as, “To what extent are your expenses driven by your children?”.
+These kinds of open-ended questions will allow for a comparison with the observational data.
+If many families are coming to grocery stores with carts full of children’s clothes and snacks, and if many families are indicating that most of their expenses center around their children, this suggests that planning decisions that help support these families may also enhance a neighbourhood’s income position.
+
+Thus, observational data and surveys can be helpful tools to find out more about individual family characteristics that may often be overlooked or lost when looking at aggregated census data or considering all Toronto neighbourhoods at once.
+
+\newpage
+
 \newpage

### 2024-12-04 plan/datasheet/datasheet_template.qmd
Updated datasheet for analysis data

---
title: "Datasheet for Toronto Neighbourhood Profiles Data"
subtitle: "A deeper look into the 2021 Canadian Census Data"
author: 
  - Julia Lee
thanks: "Code and data are available at: https://github.com/jjlee-lee/Toronto_Neighbourhood_Income.git"
date: today
date-format: long
abstract: "The following contains information about the Toronto Neighbourhoods Profile Data. Extract of the questions are provided by Gebru and colleagues (2021, https://doi.org/10.1145/3458723)"
format: pdf
number-sections: true
---

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The Neighbourhood Profiles dataset was created to gain a better understanding of the populations living in the city of Toronto. By highlighting various demographic and socio-economic characteristics such as education, income, and employment of Toronto residents, this data can be used to understand how Toronto is changing and to better serve all Toronto residents. 
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - This dataset is a collection of data from the 2021 Canadian Census conducted and provided by Statistics Canada. To organize population information for Toronto specifically, the Social Development, Finance & Administration of Open Data Toronto created this dataset. 
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The creation of the dataset is funded by the city of Toronto, and it is made accessible to everyone through the city's open data portal.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The instances that comprise the dataset represent Toronto residents and the various demographic and socio-economic characteristics they had at the time of the 2021 Census.
2. *How many instances are there in total (of each type, if appropriate)?*
	- There are over 2000 categories (e.g. education attainment, household size, etc.) of data provided for each of the 158 neighbourhoods in Toronto.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The short-form census is sent out to all households in a given region, and the long-form census is sent out to only 25% of households in a given region. With this, the 2021 Census data reflected in the Neighbourhood Profiles data can be a representative sample of all residents in a specific region.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance consists of unprocessed text in an excel file. 
5. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- Missing data values can be imputed and the aggregation of individual household characteristics or experiences can lead to incorrect assumptions about certain regions and their populations.

### 2024-12-15 plan/datasheet/datasheet_template.qmd
Updated the datasheet for the Toronto Neighbourhood Profiles data - included more information about data use, distribution, processing, and more.

@@
 date-format: long
-abstract: "The following contains information about the Toronto Neighbourhoods Profile Data. Extract of the questions are provided by Gebru and colleagues (2021, https://doi.org/10.1145/3458723)"
+abstract: "The following contains information about the Toronto Neighbourhoods Profile data used within the present analysis.
+Extract of the questions are provided by Gebru and colleagues (2021) and can be found at https://doi.org/10.1145/3458723."
 format: pdf
@@
 2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
-    - This dataset is a collection of data from the 2021 Canadian Census conducted and provided by Statistics Canada.
+    - This dataset is a collection of data from the 2021 Canadian census conducted and provided by Statistics Canada.
 To organize population information for Toronto specifically, the Social Development, Finance & Administration of Open Data Toronto created this dataset.
@@
 	- The short-form census is sent out to all households in a given region, and the long-form census is sent out to only 25% of households in a given region.
-With this, the 2021 Census data reflected in the Neighbourhood Profiles data can be a representative sample of all residents in a specific region.
+With this, the 2021 census data reflected in the Neighbourhood Profiles data can be a representative sample of all residents in a specific region.
 4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
@@
 	- Missing data values can be imputed and the aggregation of individual household characteristics or experiences can lead to incorrect assumptions about certain regions and their populations.
+
+**Collection process**
+
+1.
+*How was the data associated with each instance acquired?
+Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)?
+If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified?
+If so, please describe how.*
+	- The data for the Canadian census is collected through a combination of surveys (i.e.
+short and long form) where residents or households in Toronto answer questions related to demographic, social, economic, and cultural characteristics.
+The responses collected from the census are compared with data from prior censuses and government records to verify the data and ensure consistency along with validity within the data aggregation process.
+2. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
+	- The long form survey within the Canadian census employs a stratified systematic sampling design to collect data on 25% of households within a given region.
+For the Neighbourhoods Profiles data, 25% of private households in Toronto (who were sampled from a list of dwellings) would have received a long form survey.
+3. *Who was involved in the data collection process (for example, students, crowdworkers, contractors)?*
+	- The data collection process is completed by Statistics Canada every 5 years.
+4.
+*Over what timeframe was the data collected?
+Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)?
+If not, please describe the timeframe in which the data associated with the instances was created.*
+	- The census data was collected in 2021 and has a reference data of May 11, 2021.
+The instances that are found within the dataset are a reflection of the demographic, socio-economic, and cultural characteristics of Toronto households during 2021.
+Each instance further captures the change in these characteristics since the previous Canadian census in 2016.
+5.
+*Were any ethical review processes conducted (for example, by an institutional review board)?
+If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
+	- Statistics Canada conducts formal reviews on the data that is collected through the census to ensure data quality alongside an ethical collection and use of the data.
+Further details on their exact procedures and guidelines can be accessed through the following links:
+	- https://www12.statcan.gc.ca/census-recensement/2021/ref/98-26-0001/2020001/007-eng.cfm
+	- https://www.statcan.gc.ca/en/trust/integrity
+
+**Preprocessing/cleaning/labeling**
+
+1.
+*Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)?
+If so, please provide a description.
+If not, you may skip the remaining questions in this section.*
+	- Once the census has been completed by households and returned to Statistics Canada, a process to match electronically submitted, phoned, and mailed survey responses with dwellings and geographic areas begins.
+While responses submitted online or provided through a telephone interview are automatically registered, mailed responses need to be manually registered to the Data Operations Centre (DOC).
+A more detailed process is outlined at the following link:
+	- https://www12.statcan.gc.ca/census-recensement/2021/ref/98-306/2021001/chap3-eng.cfm
+2.
+*Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)?
+If so, please provide a link or other access point to the "raw" data.*
+	- Raw data for the Canadian census is not provided. Census data that has been aggregated across different geographical regions, like the Toronto Neighbourhood Profiles data, is available.
+
+**Uses**
+
+1. *Has the dataset been used for any tasks already? If so, please provide a description.*
+	- As census data is widely available for anyone to view and analyze, the data has been used by several individual researchers and organizations to better understand social phenomenon and the locations in which social events occur.
+2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
+	- No, as the Toronto Neighbourhoods Profile data is open data, it would be challenging to maintain a record of all the ways in which it was used and analyzed. 
+3. *What (other) tasks could the dataset be used for?*
+	- From education attainment to household size, the Canadian census includes a wide range of variables for different regions across Canada.
+With this, the Toronto Neighbourhoods Profile data could be used to inform decisions in urban planning, healthcare, education policies, and beyond.
+
+**Distribution**
+
+1. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
+	- The Toronto Neighbourhoods Profile data is open data, meaning that it is accessible to all and can be easily downloaded by users.
+With this, the dataset is widely distributed through the city of Toronto's online open data portal.
+2. *Any other comments?*
+	- It is important to note that while the dataset is open to all and for any type of analysis, every user has a responsibility to use the data in ethical and respectful ways.
+
+**Maintenance**
+
+1. *Who will be supporting/hosting/maintaining the dataset?*
+	- The dataset found on the city of Toronto's open data portal will be supported and maintained by the city along with the Social Development, Finance & Administration division (the division that published the data to the portal).
+2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
+	- The Social Development, Finance & Administration division can be contacted via email.
+An email address is provided on the open data portal's website where the dataset can be downloaded (https://open.toronto.ca/dataset/neighbourhood-profiles/).
+3.
+*Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)?
+If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
+	- As the dataset is a collection of information from the 2021 Canadian census, it will not be updated to add or remove new instances. 
+4.
+*If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)?
+If so, please describe these limits and explain how they will be enforced.*
+	- The data provided by residents and households who completed the 2021 census will be released to Library and Archives Canada (LAC) after 92 years.
+This release of information currently does not require their consent.
+For further information, please refer to the following link:
+	- https://www12.statcan.gc.ca/census-recensement/2021/ref/personalinfo-renpersonnels-eng.cfm
+5. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
+	- Older versions of the census, including information for Toronto, are maintained by the Library and Archives Canada (LAC).
+Online or microfilm copies of census records dating back to 1926 can be accessed through the LAC.
